---
title: "Results of Calculations for Fact Checking Pasture Report"
author: "Dan Rejto"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r global_options, echo=FALSE, message=FALSE}
#Enter code in console to navigate to Peak Pasture Folder before starting
#setwd("~/Google Drive File Stream/My Drive/BTI Research/Food & Farming/Peak Pasture")

#set knitr display options
library(knitr)
opts_chunk$set(fig.width=10, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE)

#opts_knit$set(root.dir = "../../Peak Pasture/data")
```

```{r load packages}
require('tidyverse')
require('ggthemes')
require('scales')
require('plotly')
require('rworldmap')
require('rlang')
require('broom')
```

```{r Load datasets, message=FALSE}
combined_country <- read_csv(file = "data/clean_data/combined_tidy.csv") #country-level dataset with combined cattle data
meat_cons_global <- read_csv(file =  "data/clean_data/meat_cons_global.csv") #global cattle, goat, mutton & pig meat consumption
milk_cons_global <- read_csv(file =  "data/clean_data/milk_cons_global.csv") #global milk consumption
meat_prod_global <- read_csv(file =  "data/clean_data/meat_prod_global.csv") #global herding animal production
milk_prod_global <- read_csv(file =  "data/clean_data/milk_prod_global.csv") #global milk production
meat_prod_global_adjusted <- read_csv(file = "data/clean_data/meat_prod_global_adjusted.csv") #global herding animal production, data adjusted
milk_prod_global_adjusted <- read_csv(file =  "data/clean_data/milk_prod_global_adjusted.csv") #global milk production, data adjusted
cattle_stocks_global <- read_csv(file =  "data/clean_data/cattle_stocks_global.csv") #global cattle stocks  #remove?
hyde <- read_csv(file =   "data/clean_data/hyde.csv") #country level pasture data from hyde model
fue <- read_csv(file = "data/raw_data/fue_wirsenius_2010.csv", skip = 2) #feed use efficiency
# I manually labelled regions in the FUE dataset by whether pasture area increased between 1961 and 2016, since the region names don't align well with the minor or major group names in our country-level dataset
fao_forecast <- read_csv(file = "data/clean_data/fao_beef_milk_consumption_forecast.csv") # FAO's 2030, 35, 40, 45, 50 beef and raw milk consumption forecasts by country
```

```{r Set themes}
#rtn

#set rules for formatting graphs
point <- format_format(big.mark = ",", decimal.mark = ".", scientific = FALSE)

# bti_theme <- theme_classic() +
#                         theme(base_size = 20)
# #theme(text = element_text(family = base_family, face = "plain", colour = "black", 
#                                             size = base_size, hjust = 0.5, vjust = 0.5, angle = 0, lineheight = 0.9)),
#                         legend.position = c(0.2, 0.8),
#                         legend.title = element_blank(),
#                         legend.text = element_text(size = rel(0.8)), 
#                         plot.title = element_text(size = rel(1.2)),
#                         axis.text = element_text(size = rel(0.8), colour = "grey50"),
#                         axis.title.x = element_text(), 
#                         axis.title.y = element_text(angle = 90)
#                         ))
theme_set(theme_gray())

income_labeller <- list("H" = "High", "UM" = "Upper Middle", "LM" = "Lower Middle", "L" = "Low")
climate_labeller <- list("H" = "Humid", "A" = "Arid", "T" = "Temperate")

wdt <- 8 #default figure output width
hgt <- 8 #default figure output height
big_wdt <- 16 #width for figures with 2 facets side by side
big_hgt <- 16 #height for figures with 2 facets on top of one another
```

```{r Functions}
# create dataframe with total % change, annual % change, and parameters for 1 variable. If a variable is passed that has multiple groups (e.g. meat production for cattle and goats), it summarizes the values across the groups
pct_chg <- function(data, col_name, first_yr, last_yr) {
    
  data <- data %>%  #summarize across multiple items and create new column with sum for each year
      group_by(year) %>%
      summarize(tot =  sum(!! sym(col_name)))
  
  last <- filter(data, year==last_yr)$tot
  first <- filter(data, year==first_yr)$tot
  diff_abs = last-first 
  diff_pct = ((last/first)-1)*100
  pct_annual = (log(last)-log(first))/(last_yr-first_yr)*100 # calculate annual percent change per http://econweb.rutgers.edu/rockoff/growthrate.htm
  
  temp <- tibble("first_year" = first_yr, 
                 "last_year" = last_yr,
                 "first_year_value" = first,
                 "last_year_value" = last,
                 "difference" = diff_abs, 
                 "percent_difference" = diff_pct,
                 "annual_percent_change" = pct_annual)
  return(temp)
}


#function for calculating change for one variable}
country_level_change_one_var <- function (x, var, first_yr, last_yr) {
  
  temp <- as.data.frame(x) %>% 
    filter(year %in% c(first_yr,last_yr)) %>% #filter to years of interest
    select(country, year, var) %>% #select variables of interest
    spread(key = year, value = (var))
  
  final <- tibble(
    temp[,1],
    temp[,2],
    temp[,3],
    (temp[,3]-temp[,2])/temp[,2],
    (temp[,3]-temp[,2])
    )
     
 vars <- c("country", 
           paste0(var, "_", first_yr),
           paste0(var, "_", last_yr),
           paste0(var, "_", "pct_chg"),
           paste0(var, "_", "abs_chg"))
 
 final <- setNames(final, vars)

  return(final)
}


#create dataframe with indexed values for 1 variable. E.g. A data frame with values 10, 20, 25, 20 would return a dataframe with values 1, 2, 2.5, 2
create_index <- function(data, base_year, col_name) {
 temp <- data %>%
  filter(year >= base_year) %>%
  group_by(year) %>%   #
  summarize(tot = sum(!! sym(col_name))) %>%   #pass col_name as a name instead of string
  with(., ave(tot, FUN = function(x) x/x[1]))  #calculate percent change from base year  
 
 return(temp)
}


#create dumbell graph showng change between two points in time
yield_dumbell <-function(x, animal, first_yr, last_yr) {
  temp <- x %>% filter(item==as.character(animal), year %in% c(first_yr,last_yr)) %>%
    spread(key = year, value=yield)
  
  colnames(temp)[3:4] <- c("first_yr", "last_yr")
  
  viz <- plot_ly(temp, color = I("gray80")) %>%
    add_segments(x = ~first_yr, xend = ~`last_yr`, y = ~country, yend = ~country, showlegend = FALSE) %>%
    add_markers(x = ~first_yr, y = ~country, name = deparse(first_yr), color = I("red")) %>%
    add_markers(x = ~last_yr, y = ~country, name = deparse(last_yr), color = I("green")) %>%
    layout(
      title = paste0("Change in ", str_to_title(as.character(animal)), " Meat Yields"),
      xaxis = list(title = "Yields (hg/animal)"),
      margin = list(l = 65)) 
  return(viz)
}
```

```{r annual pct and absolute change function}
pct <- function(x) ((x/dplyr::lag(x, n=1)-1)*100)

pct_since_1961 <- function(x) ((x/x[1]))
abs_since_1961 <- function(x) ((x-x[1]))
```

```{r Filter All Data}
full_data  <- combined_country #save for exploratory analysis in appendix
combined_country <- combined_country %>%
  filter(inc_group_16 %in% c("H", "UM", "LM", "L"), #omit small countries w/o income for plotting
         country !="Saudi Arabia") #omit because FAO omitted it in its pasture land use analysis, Ramankutty et al. discuss the FAO values as being orders of magnitude above subnational values, and because GDP data must be outlier

#earlier version of code omitted Iran: code for doing this is here
# !str_detect(country, "Iran"), #Omit Iran since its pasture suddenly drops ~17 Mha without explanation in 2005.

```

```{r Calculate annual pct changes by country}
combined_country <- combined_country %>%
  group_by(country) %>%
  mutate(ann_pct_chg_area = pct(area),
         ann_pct_chg_m_yld = pct(m_yld),
         ann_pct_chg_m_tot_cons = pct(m_cons_total),
         ann_pct_chg_m_cons_pc_cal = pct(m_cons_pc_cal),
         ann_pct_chg_pop = pct(pop),
         ann_pct_chg_d_yld = pct(d_yld),
         ann_pct_chg_d_tot_cons = pct(d_cons_total),
         ann_pct_chg_d_cons_pc_cal = pct(d_cons_pc_cal),
         ann_pct_chg_gdp_pc = pct(gdp_pc))
```

```{r Calculate pct & abs change by country since 1961}
combined_country <- combined_country %>%
  group_by(country) %>%
  mutate(pct_chg_area = pct_since_1961(area),
         pct_chg_m_tot_cons = pct_since_1961(m_cons_total),
         pct_chg_d_tot_cons = pct_since_1961(d_cons_total),
         pct_chg_m_prod = pct_since_1961(m_prod),
         pct_chg_d_prod = pct_since_1961(d_prod),
         pct_chg_m_cons_pc_cal = pct_since_1961(m_cons_pc_cal),
         pct_chg_d_cons_pc_cal = pct_since_1961(d_cons_pc_cal),
         pct_chg_m_yld = pct_since_1961(m_yld),
         pct_chg_d_yld = pct_since_1961(d_yld),
         pct_chg_pop = pct_since_1961(pop),
         pct_chg_gdp = pct_since_1961(gdp),
         pct_chg_stocks = pct_since_1961(stocks),
         pct_chg_stocking = pct_since_1961(stocks/area),
         abs_chg_area = abs_since_1961(area))
```

```{r Create variables for ease of analysis and plotting}
since_1961 <- c(1961, 1970, 1980, 1990, 2000, 2010, 2016)
since_1961_2013 <- c(1961, 1970, 1980, 1990, 2000, 2010)
since_1992 <- c(1992, 1995, 2000, 2005, 2010, 2015)
since_2000<- c(2000,2005,2010, 2015)

#hectares to billion hectares conversion
ha_to_bha <- 1/1000000000

#hectares to million hectares conversion
ha_to_mha <- 1/1000000

#metric ton-kg conversion
t_to_kg <- 1000
kg_to_t <- 1/1000
```

```{r Create groups for graphs}
#create grouping variable that recodes major minor and countryies 
rest_of_world <-  c("Central America","Caribbean", "South-Eastern Asia","Southern Asia", "Western Asia", "Eastern Asia", "Oceania", NA)
combined_country$group <- combined_country$major_group
combined_country[combined_country$minor_group %in% rest_of_world,]$group <- "Rest of World"

combined_country[combined_country$minor_group=="Former Soviet Union",]$group <- "Former Soviet Union"
combined_country[combined_country$minor_group=="Northern America",]$group <- "Northern America"
combined_country[combined_country$minor_group=="Northern America",]$group <- "Northern America"
combined_country[combined_country$minor_group=="South America",]$group <- "Rest of South America"
combined_country[combined_country$country =="China",]$group <- "China"
combined_country[combined_country$country =="Brazil",]$group <- "Brazil"
combined_country[combined_country$country =="Australia",]$group <- "Australia"

#create list of countries with largest pasture area in period 1992-2016 (Mexico was in top 10 pre USSR dissolution)
biggest_pasture_countries <- c("United States of America", "Brazil", "China", "Australia", "Kazakhstan", "Saudi Arabia", "Mongolia", "Argentina", "Russia", "South Africa") #
```

# Introduction Calculations

```{r Forecast change in Beef consumption 2012 to 2050}
beef_cons_12_50 <- fao_forecast %>% 
  filter(item == "Beef and veal") %>%
  group_by(year) %>% 
  summarize(tot = sum(tot_cons, na.rm = F)) %>% 
  spread(key = year, value = tot) %>% 
  mutate(pct_chg_12_50 = (`2050`-`2012`)/`2012`*100) %>%  
  pull(-1)  #convert last variable to vector

#round(beef_cons_12_50,2)

#percent in high-income countries vs. everywhere else
percent_beef_chg_12_50_by_income <- fao_forecast 

percent_beef_chg_12_50_by_income$region <- fct_other(percent_beef_chg_12_50_by_income$region, 
                                                     keep = "High-income", other_level = 'Developing')  #recode region to 2 groups

percent_beef_chg_12_50_by_income <- percent_beef_chg_12_50_by_income %>% 
  filter(item == "Beef and veal") %>%
  group_by(year, region) %>% 
  summarize(tot = sum(tot_cons, na.rm = F)) %>% 
  spread(key = year, value = tot) %>% 
  mutate(pct_chg_12_50 = (`2050`-`2012`)/`2012`*100)

```
Demand for beef alone is forecast to rise `r round(beef_cons_12_50,2)`% by 2050, with the greatest growth in developing countries. Growth in high-income countries is `r filter(percent_beef_chg_12_50_by_income, region == "High-income")$pct_chg_12_50`% and in all other countries (which we categorize as developing ones) is `r filter(percent_beef_chg_12_50_by_income, region == "Developing")$pct_chg_12_50`

# Chapter 1 Calculations

## Global Pasture Area Has Peaked

```{r Fig 1 Line Graph of global pasture area since 1961}
pasture_global <- combined_country %>%
  filter(is.na(area)==F) %>% 
  group_by(year) %>%
  summarize(area=sum(area)) %>%
  ungroup()

#calculate absolute and % decline from 2000
pasture_global_2000 <- pasture_global$area[pasture_global$year==2000]

pasture_global_latest <- pasture_global$area[pasture_global$year==max(pasture_global$year, na.rm=T)]

pasture_global_pct_fall_2000_16 <- (1-(pasture_global_latest/pasture_global_2000))*100    #percent difference between 2000 and latest year

pasture_global_abs_fall_2000_16 <- pasture_global_2000 - pasture_global_latest   #absolute difference from 2000 to latest year

#avg. annual rate of pasture change in 16 years since 2000
pasture_change_rate_post2000 <- -pasture_global_pct_fall_2000_16/16

#avg. annual rate of pasture change since 1984 to 2000 (16 years before 2000)
temp <- pasture_global %>% filter (year==1984|year==2000) #just get 1984 and 2000 values
pasture_change_rate_pre2000 <- (filter(temp, year==2000)[,2] / filter(temp, year==1984)[,2])/16

#percent change from 1961 to 2016
pasture_change_1961_2016 <- pct_chg(pasture_global, "area", 1961, 2016)

#percent change from 1961 to 2000
pasture_change_1961_2000 <- pct_chg(pasture_global, "area", 1961, 2000)

p <- ggplot(pasture_global) +
  geom_line(aes(x = year, y = area/1000000000)) +  #convert to billion ha 
  labs(title = paste0("Fig 1: Global Pasture Area Declined ", round(pasture_global_pct_fall_2000_16,1), "% (", round(pasture_global_abs_fall_2000_16/1000000,1), " Mha) since 2000"),  
       y = "Global Pasture Area (billion ha)", 
       x = "") +
  scale_x_continuous(breaks = since_1961) + 
  theme_hc()   #highcharts theme

#p

ggsave(filename = "outputs/fig_1_global_pasture_area_trend.pdf", plot = p)
```
- The decline of `r round(pasture_global_abs_fall_2000_16/1000000,1)` is equivalent to `r round(pasture_global_abs_fall_2000_16 / pasture_change_1961_2000$difference, 2)*100`% of the change in pasture from 1961 to 2000.  


```{r Australia change in pasture rate }
australia_og_pasture <- read_csv(file = "data/raw_data/fao_country_level_land_area.csv") %>% 
  filter(Area=="Australia", Item=="Land under perm. meadows and pastures")

australia_30_yr_rate <- filter(australia_og_pasture, Year %in% c(1985:2014)) %>% mutate(pct_chg=pct(Value)) %>% summarize(rate=mean(pct_chg, na.rm=T)) %>% as.numeric()

australia_2015_rate <- filter(australia_og_pasture, Year %in% c(2014,2015)) %>% mutate(pct_chg=pct(Value)) %>% summarize(rate=mean(pct_chg, na.rm=T)) %>% as.numeric()
```

- The 30-year rate of pasture change for Australia (from 1985 to 2014) was `r australia_30_yr_rate`%. The 2014-2015 rate of decline however was `r australia_2015_rate`%. This is about `r round(australia_2015_rate / australia_30_yr_rate,2)` times the 30 year rate. 

## Peak Pasture Is a World-Wide Trend
```{r Percent and number of countries where pasture declined }
#Calculate % of global pasture area of countries that have declined from 1961-2016 and from 2000 to 2016. Calculate # of countries that have declined too for these time periods

# total countries 2016
n_countries <- length(unique(combined_country$country[combined_country$year==2016]))

# countries where pasture declined 1961-2016
countries_down_61_16 <- combined_country %>% filter(year==2016, abs_chg_area<0) %>% select(country)

# number of countries where pasture declined 1961-2016
n_countries_down_61_16 <- nrow(countries_down_61_16)


# countries where pasture increased 1961-2000
countries_up_61_00 <- combined_country %>% filter(year==2000, abs_chg_area>0) %>%select(country)
n_countries_up_61_00 <- nrow(countries_up_61_00)

# #countries where pasture declined 2000-2016
countries_down_00_16 <- combined_country %>% select(country, year, abs_chg_area) %>% filter(year %in% c(2000,2016)) %>% spread(year, abs_chg_area) %>% filter(`2016`<`2000`) %>% select(country)

#number & percent of countries where pasture incerased 61-2000 then declined 00-16
n_countries_up_then_down <- nrow(intersect(countries_up_61_00,countries_down_00_16))
p_countries_up_then_down <- round(n_countries_up_then_down/n_countries_up_61_00, 2)*100

# #countries where pasture declined 2000-2016
n_countries_down_or_flat_00_16 <- combined_country %>% select(country, year, abs_chg_area) %>% filter(year %in% c(2000,2016)) %>% spread(year, abs_chg_area) %>% filter(`2016`<=`2000`) %>% nrow()

# % of countries where pasture declined 1961-2016
p_countries_down_61_16 <- round(n_countries_down_61_16 / n_countries, 2)*100

# % of countries where pasture declined 2000-2016
p_countries_down_or_flat_00_16 <- round(n_countries_down_or_flat_00_16 / n_countries, 2)*100

# % of pasture represented by countries where pasture declined 1961-2016
#p_area_down_61_16

# % of pasture represented by countries where pasture declined 2000-2016
#p_area_down_00_16
```
- Pasture area was flat or declining in `r p_countries_down_or_flat_00_16`% of countries between 2000 and 2016.
 
- Overall, pasture area in `r n_countries_down_61_16` countries (`r p_countries_down_61_16`% of countries in the dataset today) is now below where it was in 1961

- Altogether, since 2000, pasture area has contracted in `r n_countries_up_then_down` countries that saw pasture increase from 1961-2000. In other words, `r p_countries_up_then_down`% countries that saw pasture increased 1961-2000 saw it subsequently decrease. 

- Following table shows countries where pasture increased the most with their pasture change from 2000 to 2016
```{r countries where pasture increased most 2000-2016}
pasture_change_by_country_00_16 <- combined_country %>% select(country, year, area) %>% filter(year %in% c(2000,2016)) %>% 
  spread(key = year, value=area) %>% 
  mutate(change_00_16=`2016`-`2000`) %>%
  arrange(desc(change_00_16)) %>% 
  head(5) %>% kable(caption = "Countries in Descending Order by Pasture Change from 2000-2016")
pasture_change_by_country_00_16
```

## Peak Pasture Holds Positive Environmental Potential
```{r Forest area increase in europe}
forest_change_eu <- read_csv(file = "data/clean_data/land_area.csv") %>% #load land use data
  filter(major_group=="Europe", item=="forest", year %in% c(1990,2015)) %>%  #filter to Europe forest b/t 1990 & 2015
  mutate(year=as.factor(year)) %>% 
  group_by(year) %>% 
  summarize(forest_area=sum(area, na.rm=T))

thousand_to_million=1/1000
eu_forest_abs_chg <- (forest_change_eu[2,2] - forest_change_eu[1,2])*thousand_to_million #calculate change
eu_forest_pct_chg <- ((forest_change_eu[2,2] - forest_change_eu[1,2]) /forest_change_eu[1,2])*100 #calculate pct change
```
- In the 25 years between 1990 and 2015, the area of forest in Europe, a region that has experienced some of the most significant declines in pasture, has increased by about `r as.numeric(round(eu_forest_abs_chg,2))` million hectares (or `r as.numeric(eu_forest_pct_chg)`%).



## Productivity Gains Key to Peak Pasture
```{r Pct change values in production and consumption }
top_grazing_animals <- c("cattle", "goat", "sheep", "buffalo")

# Production of buffalo, cattle, goat, sheep MEAT 
meat_prod_change_2000 <- pct_chg(filter(meat_prod_global_adjusted, item %in% top_grazing_animals), "m_prod", 2000, 2013)

# Production of buffalo, cattle, goat, sheep MILK
milk_prod_change_2000 <- pct_chg(milk_prod_global_adjusted, "d_prod", 2000, 2013)

#From 1961 to 2013

# Production of buffalo, cattle, goat, sheep MEAT 
meat_prod_change_1961 <- pct_chg(meat_prod_global, "m_prod", 1961, 2013)

# Production of buffalo, cattle, goat, sheep MILK
milk_prod_change_1961 <- pct_chg(milk_prod_global, "d_prod", 1961, 2013)
```

- Between 2000 and 2013, aggregate production of meat and milk from cattle, buffaol, goats and sheep—the most common grazing animals—rose by `r meat_prod_change_2000$percent_difference`% and by `r milk_prod_change_2000$percent_difference`%, respectively.


```{r Calculate Decadal Global Consumption & Yield Changes}
base_year = 1961

temp <-   combined_country %>%
  filter(year>= base_year, year<=2013) %>%
  group_by(year) %>%
  summarize(m_cons_total= sum(m_cons_total, na.rm=T), #total consumption (technically food availability)
            d_cons_total= sum(d_cons_total, na.rm=T), #total consumption (technically food availability)
            m_yld= sum(m_prod, na.rm=T)/sum(m_head, na.rm=T), #tons production/animal
            d_yld= sum(d_prod, na.rm=T)/sum(d_head, na.rm=T), #tons production/animal
            m_tot_productivity = sum(m_prod, na.rm=T)/sum(area, na.rm=T), #total beef productivity of land ( tons beef/hectare total pasture)
            d_tot_productivity = sum(d_prod, na.rm=T)/sum(area, na.rm=T)) #total milk productivity of land ( tons milk/hectare total pasture)
  
decadal_cons_yld_changes <- tibble(
  "Period" = rep(c("1961-1970","1971-1980", "1981-1990","1991-2000","2001-2010", "2011-2013"),6),
  "Product" = c(rep("Beef", 18), rep("Dairy", 18)),
  "Variable" = c(rep("Consumption", 6), rep("Yield", 6), rep("Total Productivity", 6), rep("Consumption", 6), rep("Yield", 6) ,rep("Total Productivity", 6)),
  "Annual Percent Change" = c( 
    pct_chg(temp, "m_cons_total", 1961, 1970)[1,7], 
  pct_chg(temp, "m_cons_total", 1971, 1980)[1,7],
  pct_chg(temp, "m_cons_total", 1981, 1990)[1,7],
  pct_chg(temp, "m_cons_total", 1991, 2000)[1,7],
  pct_chg(temp, "m_cons_total", 2001, 2010)[1,7],
  pct_chg(temp, "m_cons_total", 2011, 2013)[1,7],
  
  pct_chg(temp, "m_yld", 1961, 1970)[1,7], 
  pct_chg(temp, "m_yld", 1971, 1980)[1,7],
  pct_chg(temp, "m_yld", 1981, 1990)[1,7],
  pct_chg(temp, "m_yld", 1991, 2000)[1,7],
  pct_chg(temp, "m_yld", 2001, 2010)[1,7],
  pct_chg(temp, "m_yld", 2011, 2013)[1,7],
  
  pct_chg(temp, "m_tot_productivity", 1961, 1970)[1,7], 
  pct_chg(temp, "m_tot_productivity", 1971, 1980)[1,7],
  pct_chg(temp, "m_tot_productivity", 1981, 1990)[1,7],
  pct_chg(temp, "m_tot_productivity", 1991, 2000)[1,7],
  pct_chg(temp, "m_tot_productivity", 2001, 2010)[1,7],
  pct_chg(temp, "m_tot_productivity", 2011, 2013)[1,7],
  
  pct_chg(temp, "d_cons_total", 1961, 1970)[1,7], 
  pct_chg(temp, "d_cons_total", 1971, 1980)[1,7],
  pct_chg(temp, "d_cons_total", 1981, 1990)[1,7],
  pct_chg(temp, "d_cons_total", 1991, 2000)[1,7],
  pct_chg(temp, "d_cons_total", 2001, 2010)[1,7],
  pct_chg(temp, "d_cons_total", 2011, 2013)[1,7],
  
  pct_chg(temp, "d_yld", 1961, 1970)[1,7], 
  pct_chg(temp, "d_yld", 1971, 1980)[1,7],
  pct_chg(temp, "d_yld", 1981, 1990)[1,7],
  pct_chg(temp, "d_yld", 1991, 2000)[1,7],
  pct_chg(temp, "d_yld", 2001, 2010)[1,7],
  pct_chg(temp, "d_yld", 2011, 2013)[1,7],
  
  pct_chg(temp, "d_tot_productivity", 1961, 1970)[1,7], 
  pct_chg(temp, "d_tot_productivity", 1971, 1980)[1,7],
  pct_chg(temp, "d_tot_productivity", 1981, 1990)[1,7],
  pct_chg(temp, "d_tot_productivity", 1991, 2000)[1,7],
  pct_chg(temp, "d_tot_productivity", 2001, 2010)[1,7],
  pct_chg(temp, "d_tot_productivity", 2011, 2013)[1,7])
  )
```

```{r Fig 7 Smooth Graph of Annual Pct change in beef and cattle milk demand}
#calculate annual % change for each meat item
pd1 <- meat_cons_global %>% 
  select(year, cons_total, item) %>% #just calculate total consumption change
  group_by(item) %>%  mutate(pct_chg = (cons_total/dplyr::lag(cons_total)-1)*100) #annual pct change by animal

#calculate annual % change for each dairy
pd2 <- milk_cons_global %>% 
  select(year, cons_total, item) %>% #just calculate total consumption change
  group_by(item) %>%  mutate(pct_chg = (cons_total/dplyr::lag(cons_total)-1)*100) #annual pct change by animal

pd <- bind_rows(pd1, pd2)

p <- pd %>% 
  filter(item %in% c("Bovine Meat", "Milk - Excluding Butter")) %>%  #just graph for beef &milk
  ggplot(aes(x=year, y=pct_chg/100, color=item))+
  geom_smooth(method = "loess", se = F) + #show loess line but hide standard error bar
  geom_point(alpha=.2) + 
  labs(y = paste0("Annual Percent Change"),
       x = "") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = since_1961) +
    theme_hc()+
  theme(legend.position = "none")+
  facet_wrap(item~.)
#p

ggsave(filename = "outputs/fig_7_annual_beef_milk_consumption_change.pdf", plot = p)
```

```{r Calculate pct change in global cattle meat and milk yields, production and consumption}
base_year = 1961
global_indices <- combined_country %>%
  filter(minor_group != "Former Soviet Union") %>%  #omit USSR and former USSR
  #filter(year>= base_year, year<=2013) %>%
  group_by(year) %>%
  summarize(m_yld = sum(m_prod, na.rm=T)/sum(m_head, na.rm=T),  #beef yield
            m_cons = sum(m_cons_total, na.rm=T),  # beef consumption total
            m_prod = sum(m_prod, na.rm=T), #beef production
            d_yld = sum(d_prod, na.rm=T)/sum(m_head, na.rm=T),  #milk yield
            d_cons = sum(d_cons_total, na.rm=T),  # milk consumption total
            d_prod = sum(d_prod, na.rm=T), #milk production
            area = sum(area, na.rm=T),  # pasture area
            stocks = sum(stocks, na.rm=T)) %>% # stocks %>%
  mutate(stocking = stocks/area) %>% #stocking density proxy
  dplyr::ungroup()%>%
  mutate(pct_chg_m_yld = pct_since_1961(m_yld),
         pct_chg_d_yld = pct_since_1961(d_yld),
         pct_chg_m_prod = pct_since_1961(m_prod),
         pct_chg_d_prod = pct_since_1961(d_prod),
         pct_chg_area = pct_since_1961(area),
         pct_chg_stocking = pct_since_1961(stocking),
         pct_chg_stocks = pct_since_1961(stocks))

global_m_yld_pct_chg_61_13 <- round( (filter(global_indices, year==2013)$pct_chg_m_yld-1)*100, 2)
global_d_yld_pct_chg_61_13 <- round( (filter(global_indices, year==2013)$pct_chg_d_yld-1)*100, 2)
```

- Cattle, the most dominant user of pasture globally, saw meat and milk yields grow `r global_m_yld_pct_chg_61_13`% and `r global_d_yld_pct_chg_61_13`% since 1961.

```{r Fig 6 - Line Graph of pct chnage in global pasture, beef production, and cattle milk production}
p <- global_indices %>%
  select(year, pct_chg_m_prod, pct_chg_d_prod, pct_chg_area, pct_chg_stocks) %>%
  gather(key = "Variable", value = "value",  -year) %>%
  ggplot()+
    geom_line(aes(x=year, y=(value-1), color=Variable))+
  scale_x_continuous(breaks = since_1961, limits = c(base_year, 2013)) +
  labs(title = "Production Has Increased While Pasture Area Has Plateaued",
       y = paste0("Percent Change Since ", base_year),
       x = "",
       caption="Omitting Former Soviet Union") +
  scale_y_continuous(labels = scales::percent_format(), limits=c(0,NA))+
  scale_color_discrete(name = element_blank(),
                       breaks=c("pct_chg_area", "pct_chg_d_prod", "pct_chg_m_prod", "pct_chg_stocks"),
                       labels=c("Pasture Area", "Milk Production", "Beef Production", "Number of Cattle")) +
  theme_hc()
#p
ggsave(filename = "outputs/fig_6_beef_milk_global_production_pct_chg.pdf", plot = p, width = wdt, height=hgt)
```

```{r Calculate change in  meat & milk yields and stocking by geographic group}
# set baseline year
baseline_year = 1961

regional_indices <- combined_country %>%
  filter(group !="Former Soviet Union") %>% 
  group_by(year, group) %>%
  summarize(m_yld = sum(m_prod, na.rm=T)/sum(m_head, na.rm=T),  #regional beef yield
            d_yld = sum(d_prod, na.rm=T)/sum(d_head, na.rm=T),  #regional milk yield
            m_cons = sum(m_cons_total, na.rm=T),  #regional beef consumption total
            d_cons = sum(d_cons_total, na.rm=T),  #regional dairy consumption total
            area = sum(area, na.rm=T),  #regional pasture area
            stocks = sum(stocks, na.rm=T)) %>% #regional cattle stocks 
  mutate(stocking = stocks/area) %>% #regional cattle stocking density
  dplyr::ungroup()%>%
  group_by(group) %>%
  mutate(pct_chg_m_yld = pct_since_1961(m_yld),  #calculate percent change since 1961 for each year for each variable by region group
         pct_chg_d_yld =pct_since_1961(d_yld),
         pct_chg_m_cons = pct_since_1961(m_cons),
         pct_chg_d_cons = pct_since_1961(d_cons),
         pct_chg_area = pct_since_1961(area),
         pct_chg_stocking = pct_since_1961(stocking))

australia_beef_yld<- filter(regional_indices, year==2013, group=="Australia")$m_yld
africa_beef_yld_pct_lower_aus <- round(1 - (filter(regional_indices, year==2013, group=="Africa")$m_yld / australia_beef_yld),2)*100
brazil_beef_yld_pct_lower_aus <- round(1- (filter(regional_indices, year==2013, group=="Brazil")$m_yld / australia_beef_yld), 2)*100
china_yld_pct_lower_aus <- round(1 - (filter(regional_indices, year==2013, group=="China")$m_yld / australia_beef_yld), 2)*100
```

- For instance, beef yields in Africa, Brazil, and China are all substantially lower than in higher-income Australia— `r africa_beef_yld_pct_lower_aus`%, `r brazil_beef_yld_pct_lower_aus`% and `r china_yld_pct_lower_aus` % less, respectively.


```{r Fig A4 - Graph change in  meat & milk yields and stocking by group}
p <- regional_indices %>%
  filter(year==2013) %>% #filter to just see % change from 1961-2013
  select(year, group, pct_chg_d_yld, pct_chg_m_yld) %>% #select just yields and stocking to graph  # can add  pct_chg_stocking
  gather(key = variable, value = "value", -year, -group) %>% #gather variables into one column for easy graphing
  ggplot(aes(x= reorder(group, -value), y=value-1))+
    geom_col(aes(fill=variable, group=variable), position="dodge") +
    labs(title="Change in Dairy and Beef Yields by Region (1961-2016)",
         x=element_blank(),
         y="Percent Increase Since 1961",
         caption = "Omitting Former Soviet Union") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    #scale_x_continuous(breaks = since_1961) +
    theme(legend.title = element_blank(), legend.position = "none")+
    scale_fill_hc(name=element_blank(),
                         breaks=c("pct_chg_d_yld", "pct_chg_m_yld", "pct_chg_stocking"),
                         labels=c("Dairy Yields", "Beef Yields", "Cattle Stocks per Pasture Area"))+
    theme_hc()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#p
ggsave(filename = "outputs/fig_a4_change_in_yld_by_geographic_group.pdf", p, width=wdt, height=hgt)
```

- Globally, the number of cattle stocks per hectare has increased `r round(filter(global_indices, year==2016)$pct_chg_stocking-1,2)*100`% since 1961


```{r Calculation of Extra land needed if efficiency were at 1961 levels }
#calcualte global meat and dairy producgtion per hectare for 1961 and 2013
m_prod_per_ha_1961 <- combined_country %>% filter(year==1961) %>% group_by(year) %>% summarize(tot = sum(m_prod, na.rm=T)/sum(area, na.rm=T))%>% select(tot)
d_prod_per_ha_1961 <- combined_country %>% filter(year==1961) %>% group_by(year) %>% summarize(tot = sum(d_prod, na.rm=T)/sum(area, na.rm=T))%>% select(tot)
m_prod_per_ha_2013 <- combined_country %>% filter(year==2013) %>% group_by(year) %>% summarize(tot = sum(m_prod, na.rm=T)/sum(area, na.rm=T))%>% select(tot)
d_prod_per_ha_2013 <- combined_country %>% filter(year==2013) %>% group_by(year) %>% summarize(tot = sum(d_prod, na.rm=T)/sum(area, na.rm=T))%>% select(tot)

#calculate total meat and dairy production in 2013
m_production_2013 <- combined_country %>% filter(year==2013) %>% group_by(year) %>% summarize(tot=sum(m_prod, na.rm=T))%>% select(tot)
d_production_2013 <- combined_country %>% filter(year==2013) %>% group_by(year) %>% summarize(tot=sum(d_prod, na.rm=T)) %>% select(tot)

#pct difference in 1961 and 2013 production per hectare
beef_pct_chg_prod_per_ha_61_13 <- ((m_prod_per_ha_2013-m_prod_per_ha_1961)/m_prod_per_ha_1961) %>% pull()
milk_pct_chg_prod_per_ha_61_13 <- ((d_prod_per_ha_2013-d_prod_per_ha_1961)/d_prod_per_ha_1961) %>% pull()

#calculate extra land needed to produce 2013 meat and dairy with 1961 production/hectare levels
m_extra_land <- (m_production_2013 / m_prod_per_ha_1961) - (m_production_2013 / m_prod_per_ha_2013) #extra land assuming its all for beef
d_extra_land <- (d_production_2013 / d_prod_per_ha_1961) - (d_production_2013 / d_prod_per_ha_2013) #extra land assuming its all for milk

min_extra_land <- min(m_extra_land, d_extra_land) #get minimum of the two values above to give a conservative estimate

total_pasture_2013 <- pasture_global[pasture_global$year==2013,]$area
min_pct_extra_land <- min_extra_land/total_pasture_2013
```

- In 2013, approximately `r round(beef_pct_chg_prod_per_ha_61_13, 2)*100`% and `r round(milk_pct_chg_prod_per_ha_61_13, 2)*100`%   more cattle meat and dairy was produced for each hectare of pasture than in 1961.

- If the efficiency of the global cattle system was the same in 2016 as it was in 1961, about `r round(min_extra_land*ha_to_bha,2)` billion more hectares of pasture would be required to support the same level of production.
